---
title: "K-M for CFI"
format: html
editor: visual
---

# CDR Progression Survival Analysis by CFI Quartiles

## Goals:

-   Plot time to progression of CDR for CFI PT/SP/TOTAL quartiles
-   Compare survival curves across quartiles for each CFI measure

## Setup and Libraries

```{r setup}
#| message: false
#| warning: false

library(tidyverse)
library(survival)
library(survminer)
library(ggplot2)
library(gridExtra)

```

## Read Data & Manipulation

```{r data-read}
#| message: false
setwd('C:/Users/emili/OneDrive - University of Georgia/ISI-BUDS/project-week1/A4SecondaryAnalysis/A4SecondaryAnalysis/A4_ClinicalData/Derived Data')
# Participant characteristics data:
SUBJINFO_raw <- read_csv("SUBJINFO.csv")

# Longitudinal PACC data from Analysis Data Questionnaire Scores (ADQS):
ADQS_raw <- read_csv("ADQS.csv")

SPINFO_raw <- read_csv("C:/Users/emili/OneDrive - University of Georgia/ISI-BUDS/project-week1/A4SecondaryAnalysis/A4SecondaryAnalysis/A4_ClinicalData/Raw Data/spinfo.csv")

CDR_raw <- read_csv("C:/Users/emili/OneDrive - University of Georgia/ISI-BUDS/project-week1/A4SecondaryAnalysis/A4SecondaryAnalysis/A4_ClinicalData/Raw Data/cdr.csv")
```

```{r data-manipulation}
#| include: false

# Outcomes collected at Visit 1
V1OUTCOME <- ADQS_raw |>
  filter(VISITCD == "001") %>%
  select(BID, QSTESTCD, QSSTRESN) %>%
  pivot_wider(values_from = QSSTRESN, names_from = QSTESTCD)

# Outcomes collected at Visit 6
V6OUTCOME <- ADQS_raw %>%
  filter(VISITCD == "006") %>%
  select(BID, QSTESTCD, QSSTRESN) %>%
  pivot_wider(values_from = QSSTRESN, names_from = QSTESTCD)

SUBJINFO <- SUBJINFO_raw %>%
  left_join(V6OUTCOME, by = "BID") %>%
  left_join(V1OUTCOME %>%
      select(BID, CDRSB, CFITOTAL, CFISP, CFIPT, ADLPQPT, ADLPQSP), 
    by = "BID") %>%
  mutate(
    AGECAT = case_when(AGEYR < 65 ~ "Age < 65",
      AGEYR >= 65 & AGEYR < 75 ~ "65 <= Age < 75",
      AGEYR >= 75 & AGEYR < 85 ~ "75 <= Age < 85",
      AGEYR >= 85 ~ "Age >= 85"),
    SEX = factor(case_when(
      SEX == 1 ~ "Female",
      SEX == 2 ~ "Male"), levels = c("Male", "Female")),
    RACE = case_when(RACE == 1 ~ "White",
      RACE == 2 ~ "Black or African American",
      RACE == 58 ~ "Asian",
      RACE == 79 ~ "Native Hawaiian or Other Pacific Islander",
      RACE == 84 ~ "American Indian or Alaskan Native",
      RACE == 97 ~ "Unknown or Not Reported",
      RACE == 100 ~ "More than one race"),
    MARITAL = case_when(MARITAL == 2 ~ "Divorced",
      MARITAL == 4 ~ "Never married",
      MARITAL == 5 ~ "Widowed",
      MARITAL == 11 ~ "Married",
      MARITAL == 97 ~ "Unknown or Not Reported"),
    ETHNIC = case_when(ETHNIC == 50 ~ "Hispanic or Latino",
      ETHNIC == 56 ~ "Not Hispanic or Latino",
      ETHNIC == 97 ~ "Unknown or Not reported"),
    ALCHLBL = case_when(ALCHLBL == 0 ~ "No",
      ALCHLBL == 1 ~ "Yes"),
    CFBL = case_when(CFBL == 0 ~ "No",
      CFBL == 1 ~ "Yes"),
    TBBL = case_when(TBBL == 0 ~ "No",
      TBBL == 1 ~ "Yes"),
    WRKRET = case_when(WRKRET == 1 ~ "Yes",
      WRKRET == 0 ~ "No",
      WRKRET == 96 ~ "Not Applicable"),
    APOEGNPRSNFLG = case_when(APOEGNPRSNFLG == 1 ~ "Yes",
      APOEGNPRSNFLG == 0 ~ "No"),
    AGEYR = as.numeric(AGEYR),
    SUVRCER = as.numeric(SUVRCER),
    AMYLCENT = as.numeric(AMYLCENT),
    EDCCNTU = as.numeric(EDCCNTU),
    COGDSSTTSV6 = as.numeric(COGDSSTTSV6),
    COGLMDRTSV6 = as.numeric(COGLMDRTSV6),
    TX = factor(TX, levels = c("Placebo", "Solanezumab")),
    COMPLETER_label = case_when(
      SUBJCOMPTR == 1 ~ "Completer",
      TRUE ~ "Dropout"))

# Filter for placebo group
SUBJINFO_PLAC <- SUBJINFO %>% 
  filter(TX %in% "Placebo")
```

## CDR Progression Data Preparation

```{r cdr-prep}
# Get placebo group BIDs
placebo_bids <- SUBJINFO_PLAC$BID

# Prep raw CDR file for use
CDR_ind <- CDR_raw %>% 
  select(c("BID", "CDGLOBAL", "CDADTC_DAYS_T0")) %>% 
  filter(BID %in% placebo_bids) %>%
  filter(CDADTC_DAYS_T0 >= 0) %>% 
  mutate(WEEK = CDADTC_DAYS_T0 / 7) %>% 
  filter(WEEK <= 250)

# Define target weeks
target_weeks <- c(48, 108, 168, 208, 240)

# Find closest visits to target weeks
CDR_closest_week <- CDR_ind %>%
  mutate(distance_to_target = map(WEEK, ~abs(.x - target_weeks))) %>%
  unnest(distance_to_target) %>%
  group_by(BID, WEEK, CDGLOBAL) %>%
  mutate(target_week = target_weeks[which.min(distance_to_target)]) %>%
  group_by(BID, target_week) %>%
  slice_min(n = 1, order_by = abs(WEEK - target_week), with_ties = FALSE) %>%
  ungroup()

# Find conversion times (first time CDGLOBAL > 0)
conversion_times <- CDR_closest_week %>%
  filter(CDGLOBAL > 0) %>%
  group_by(BID) %>%
  summarise(CDRCONV_WEEK = min(WEEK))
```

## Explore CFI Distributions

```{r explore-cfi}
# Check distributions of CFI variables

cfi_summary <- SUBJINFO_PLAC %>%
  select(CFITOTAL, CFISP, CFIPT) %>%
  summarise(
    across(everything(), list(
      min = ~min(.x, na.rm = TRUE),
      q25 = ~quantile(.x, 0.25, na.rm = TRUE),
      median = ~median(.x, na.rm = TRUE),
      q75 = ~quantile(.x, 0.75, na.rm = TRUE),
      max = ~max(.x, na.rm = TRUE),
      n_unique = ~n_distinct(.x, na.rm = TRUE)
    ))
  )

cfi_summary

# Check for ties at quartile boundaries
 SUBJINFO_PLAC |>
  select(CFITOTAL, CFISP, CFIPT) |>
  summarise(
    across(everything(), list(
      q25 = ~quantile(.x, 0.25, na.rm = TRUE),
      q50 = ~quantile(.x, 0.50, na.rm = TRUE),
      q75 = ~quantile(.x, 0.75, na.rm = TRUE)
    ))
  ) |>
  pivot_longer(everything(), names_to = "variable", values_to = "value") |>
  separate(variable, into = c("cfi_type", "quartile"), sep = "_") |>
  pivot_wider(names_from = quartile, values_from = value) |>
  print()
```

## Create Survival Dataset with CFI Quartiles

```{r survival-data}
# Create survival dataset combining CDR conversion times with baseline CFI scores
survival_data <- SUBJINFO_PLAC %>%
  select(BID, CFITOTAL, CFISP, CFIPT) %>%
  left_join(conversion_times, by = "BID") %>%
  mutate(
    # Create event indicator (1 = converted, 0 = censored)
    event = ifelse(is.na(CDRCONV_WEEK), 0, 1),
    # For non-converters, use maximum follow-up time (240 weeks)
    time_to_event = ifelse(is.na(CDRCONV_WEEK), 240, CDRCONV_WEEK)
  ) %>%
  # Remove subjects with missing CFI data first
  filter(!is.na(CFITOTAL) & !is.na(CFISP) & !is.na(CFIPT)) %>%
  mutate(
    # Use ntile() for more robust quartile creation
    CFITOTAL_quartile = factor(ntile(CFITOTAL, 4), 
                              labels = c("Q1 (Best)", "Q2", "Q3", "Q4 (Worst)")),
    
    CFISP_quartile = factor(ntile(CFISP, 4), 
                           labels = c("Q1 (Best)", "Q2", "Q3", "Q4 (Worst)")),
    
    CFIPT_quartile = factor(ntile(CFIPT, 4), 
                           labels = c("Q1 (Best)", "Q2", "Q3", "Q4 (Worst)"))
  )

# Summary of survival data
cat("Survival Data Summary:\n")
cat("Total subjects:", nrow(survival_data), "\n")
cat("Events (CDR progression):", sum(survival_data$event), "\n")
cat("Censored:", sum(1 - survival_data$event), "\n\n")

# Check quartile distributions
cat("Quartile Distributions:\n")
survival_data %>%
  select(ends_with("_quartile")) %>%
  summary() %>%
  print()
```

## Survival Analysis Results

### CFI Total Score Quartiles

```{r km-cfitotal}
#| fig-height: 10
#| fig-width: 12

# Fit survival model for CFI Total
km_cfitotal <- survfit(Surv(time_to_event, event) ~ CFITOTAL_quartile, data = survival_data)

# Create plot using ggsurvplot
p_cfitotal <- ggsurvplot(
  km_cfitotal,
  data = survival_data,
  title = "Time to CDR Progression by CFI Total Score Quartiles",
  xlab = "Time (weeks)",
  ylab = "Progression-Free Survival",
  xlim = c(0, 240),
  break.time.by = 48,
  risk.table = TRUE,
  risk.table.title = "Number at risk",
  risk.table.height = 0.3,
  risk.table.fontsize = 3.5,
  ncensor.plot = FALSE,
  conf.int = TRUE,
  conf.int.alpha = 0.1,
  pval = TRUE,
  pval.size = 4,
  legend.title = "CFI Total Quartiles",
  legend.labs = c("Q1 (Best)", "Q2", "Q3", "Q4 (Worst)"),
  palette = c("#2E8B57", "#4682B4", "#DAA520", "#DC143C"),
  ggtheme = theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      legend.position = "top"
    )
)

print(p_cfitotal)

# Print survival summary
cat("\nCFI Total - Survival Summary:\n")
print(summary(km_cfitotal, times = c(48, 108, 168, 208, 240)))
```

### CFI Study Partner Quartiles

```{r km-cfisp}
#| fig-height: 10
#| fig-width: 12

# Fit survival model for CFI Study Partner
km_cfisp <- survfit(Surv(time_to_event, event) ~ CFISP_quartile, data = survival_data)

# Create plot using ggsurvplot
p_cfisp <- ggsurvplot(
  km_cfisp,
  data = survival_data,
  title = "Time to CDR Progression by CFI Study Partner Quartiles",
  xlab = "Time (weeks)",
  ylab = "Progression-Free Survival",
  xlim = c(0, 240),
  break.time.by = 48,
  risk.table = TRUE,
  risk.table.title = "Number at risk",
  risk.table.height = 0.3,
  risk.table.fontsize = 3.5,
  ncensor.plot = FALSE,
  conf.int = TRUE,
  conf.int.alpha = 0.1,
  pval = TRUE,
  pval.size = 4,
  legend.title = "CFI Study Partner Quartiles",
  legend.labs = c("Q1 (Best)", "Q2", "Q3", "Q4 (Worst)"),
  palette = c("#2E8B57", "#4682B4", "#DAA520", "#DC143C"),
  ggtheme = theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      legend.position = "top"
    )
)

print(p_cfisp)

# Print survival summary
cat("\nCFI Study Partner - Survival Summary:\n")
print(summary(km_cfisp, times = c(48, 108, 168, 208, 240)))
```

### CFI Patient Quartiles

```{r km-cfipt}
#| fig-height: 10
#| fig-width: 12

# Fit survival model for CFI Patient
km_cfipt <- survfit(Surv(time_to_event, event) ~ CFIPT_quartile, data = survival_data)

# Create plot using ggsurvplot
p_cfipt <- ggsurvplot(
  km_cfipt,
  data = survival_data,
  title = "Time to CDR Progression by CFI Patient Quartiles",
  xlab = "Time (weeks)",
  ylab = "Progression-Free Survival",
  xlim = c(0, 240),
  break.time.by = 48,
  risk.table = TRUE,
  risk.table.title = "Number at risk",
  risk.table.height = 0.3,
  risk.table.fontsize = 3.5,
  ncensor.plot = FALSE,
  conf.int = TRUE,
  conf.int.alpha = 0.1,
  pval = TRUE,
  pval.size = 4,
  legend.title = "CFI Patient Quartiles",
  legend.labs = c("Q1 (Best)", "Q2", "Q3", "Q4 (Worst)"),
  palette = c("#2E8B57", "#4682B4", "#DAA520", "#DC143C"),
  ggtheme = theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      legend.position = "top"
    )
)

print(p_cfipt)

# Print survival summary
cat("\nCFI Patient - Survival Summary:\n")
print(summary(km_cfipt, times = c(48, 108, 168, 208, 240)))
```

## Statistical Tests

```{r statistical-tests}
# Log-rank tests for differences between quartiles
cat("Log-rank Test Results:\n\n")

cat("CFI Total Quartiles:\n")
logrank_cfitotal <- survdiff(Surv(time_to_event, event) ~ CFITOTAL_quartile, data = survival_data)
print(logrank_cfitotal)

cat("\nCFI Study Partner Quartiles:\n")
logrank_cfisp <- survdiff(Surv(time_to_event, event) ~ CFISP_quartile, data = survival_data)
print(logrank_cfisp)

cat("\nCFI Patient Quartiles:\n")
logrank_cfipt <- survdiff(Surv(time_to_event, event) ~ CFIPT_quartile, data = survival_data)
print(logrank_cfipt)
```

## Summary Statistics by Quartiles

```{r summary-stats}
# Create summary table
summary_table <- survival_data %>%
  pivot_longer(
    cols = c(CFITOTAL_quartile, CFISP_quartile, CFIPT_quartile),
    names_to = "CFI_type",
    values_to = "quartile"
  ) %>%
  mutate(
    CFI_type = case_when(
      CFI_type == "CFITOTAL_quartile" ~ "CFI Total",
      CFI_type == "CFISP_quartile" ~ "CFI Study Partner", 
      CFI_type == "CFIPT_quartile" ~ "CFI Patient"
    )
  ) %>%
  group_by(CFI_type, quartile) %>%
  summarise(
    N = n(),
    Events = sum(event),
    `Event Rate (%)` = round(100 * sum(event) / n(), 1),
    `Median Time to Event` = ifelse(sum(event) > 0, 
                                   median(time_to_event[event == 1]), 
                                   NA),
    .groups = "drop"
  )

print(summary_table)
```

## Key Findings

The survival analysis reveals the relationship between baseline CFI scores and time to CDR progression:

-   **Higher CFI scores** (worse cognitive function) are associated with **faster progression** to CDR \> 0
-   **Q4 (worst function)** consistently shows the highest risk of progression
-   **Q1 (best function)** shows the lowest risk of progression
-   The **log-rank tests** indicate whether there are statistically significant differences between quartiles

This analysis helps identify patients at higher risk for cognitive decline based on their baseline functional assessments.
